\ifdefined\MAINTEXT \else Supplemental Material: \\ \fi Low-cost orbital-based linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems

### 2018-03-06 almo-langevin.tex
title

\title{
\ifdefined\MAINTEXT
\else
Supplemental Material: \\
\fi
Low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems
%RZK: without one-electron density matrix
}

@@

The new AIMD method utilizes a recently developed LS DFT~\cite{a:almo-ls} based on absolutely localized molecular orbitals (\mbox{ALMOs}). 
Unlike delocalized KS orbitals, each \mbox{ALMO} has its own \emph{localization center} and a predefined \emph{localization radius} $R_{c}$ that typically includes nearby atoms or molecules~\cite{a:stoll,a:almo-ls}. 
In the current implementation, a localization center is defined as a set of all Gaussian atomic orbitals of one molecule. 
However the approach can use other local and nonlocal basis sets~\cite{a:galli_loc, Lin2012}. 
The key feature of ALMO DFT is that its one-electron wavefunctions are constructed in a two-stage self-consistent-field (SCF) procedure~\cite{a:almo-ls} to circumvent the problem of the sluggish variational optimization emphasized above. 
In the first stage, ALMOs are constrained to their localization centers~\cite{a:khal} whereas, in the second stage, ALMOs are relaxed to allow delocalization onto the neighbor molecules within their localization radius $R_{c}$. 
To achieve a robust optimization in the problematic second stage, it is important to keep the delocalization component of the trial wavefunction orthogonal to the fixed orbitals obtained in the first stage. 
For mathematical details, see ALMO SCF method in Ref.~\citenum{a:almo-ls}.

@@

ALMO AIMD was implemented in CP2K, an open source materials modeling package~\cite{www:cp2k}. 
Accuracy and efficiency of ALMO AIMD was tested using liquid water as an example. 
This system is challenging because intermolecular electron delocalization is a critical component of hydrogen bonding and must be described correctly to reproduce static and dynamical properties of liquid water. 
%To calculate properties of water at ambient conditions, 
A periodic cell containing 125 molecules was simulated for $\SI{30}{\ps}$ at $T=\SI{298}{\K}$ and a constant density of $\SI{1.01}{\g\cdot\cm^{-3}}$. 
Ricci-Ciccotti algorithm~\cite{Ricci2003} was used to integrate the Langevin equation. We found that $\gamma = \SI{e-3}{\per\fs}$ is large enough to thermostat the system efficiently and small enough not to significantly affect dynamical properties of liquid water.
%
In the dual Gaussian and plane-wave scheme implemented in CP2K~\cite{a:quickstep}, the TZV2P basis set was used to represent molecular orbitals, and a plane-wave cutoff of $\SI{320}{Ry}$ used to represent electron density. 
The XC energy was approximated using the dispersion-corrected PBE functional~\cite{a:pbe,Grimme2010}. 
Separable norm-conserving pseudopotentials were used~\cite{a:hgh} and the Brillouin zone was sampled at the $\Gamma$-point. 
%to describe the interactions between the valence electrons and the ionic cores. 
%
%Although not a required component of ALMO AIMD, 
The predictor of the Kolafa scheme~\cite{Kolafa2003} was adopted to localized orbitals~\cite{a:2ndcpmd} to generate a highly accurate initial ALMOs in both SCF stages, which can be brought close to the ground state with just a few SCF steps of the robust two-stage optimization procedure. 
%RZK: The time-reversibility of ASPC is not a critical reqirement of the proposed method.

>>>>

\title{
\ifdefined\MAINTEXT
\else
Supplemental Material: \\
\fi
Low-cost orbital-based linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems
}

@@

The new AIMD method utilizes a recently developed LS DFT~\cite{a:almo-ls} based on absolutely localized molecular orbitals (\mbox{ALMOs}). 
Unlike delocalized KS orbitals, each \mbox{ALMO} has its own \emph{localization center} and a predefined \emph{localization radius} $R_{c}$ that typically includes nearby atoms or molecules~\cite{a:stoll,a:almo-ls}. 
In the current implementation, a localization center is defined as a set of all Gaussian atomic orbitals of one molecule. 
However the approach can use other local and nonlocal basis sets~\cite{a:ls-galli-parrinello-1992, Lin2012}. 
The key feature of ALMO DFT is that its one-electron wavefunctions are constructed in a two-stage self-consistent-field (SCF) procedure~\cite{a:almo-ls} to circumvent the problem of the sluggish variational optimization emphasized above. 
In the first stage, ALMOs are constrained to their localization centers~\cite{a:khal} whereas, in the second stage, ALMOs are relaxed to allow delocalization onto the neighbor molecules within their localization radius $R_{c}$. 
To achieve a robust optimization in the problematic second stage, it is important to keep the delocalization component of the trial wavefunction orthogonal to the fixed orbitals obtained in the first stage. 
For mathematical details, see ALMO SCF method in Ref.~\citenum{a:almo-ls}.

@@

ALMO AIMD was implemented in CP2K, an open source materials modeling package~\cite{www:cp2k}. 
Accuracy and efficiency of ALMO AIMD was tested using liquid water as an example. 
This system is challenging because intermolecular electron delocalization is a critical component of hydrogen bonding and must be described correctly to reproduce static and dynamical properties of liquid water. 
%To calculate properties of water at ambient conditions, 
A periodic cell containing 125 molecules was simulated for $\SI{30}{\ps}$ at $T=\SI{298}{\K}$ and a constant density of $\SI{1.01}{\g\cdot\cm^{-3}}$. 
Ricci-Ciccotti algorithm~\cite{Ricci2003} was used to integrate the Langevin equation. We found that $\gamma = \SI{e-3}{\per\fs}$ is large enough to thermostat the system efficiently and small enough not to significantly affect dynamical properties of liquid water.
%
In the dual Gaussian and plane-wave scheme implemented in CP2K~\cite{a:quickstep}, the TZV2P basis set was used to represent molecular orbitals, and a plane-wave cutoff of $\SI{320}{Ry}$ used to represent electron density. 
The XC energy was approximated using the dispersion-corrected PBE functional~\cite{a:PBEfunctional,Grimme2010}. 
Separable norm-conserving pseudopotentials were used~\cite{a:hgh} and the Brillouin zone was sampled at the $\Gamma$-point. 
%to describe the interactions between the valence electrons and the ionic cores. 
%
%Although not a required component of ALMO AIMD, 
The predictor of the Kolafa scheme~\cite{Kolafa2003} was adopted to localized orbitals~\cite{a:2ndcpmd} to generate a highly accurate initial ALMOs in both SCF stages, which can be brought close to the ground state with just a few SCF steps of the robust two-stage optimization procedure. 
%RZK: The time-reversibility of ASPC is not a critical reqirement of the proposed method.
